---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# tabnet

<!-- badges: start -->

[![R build status](https://github.com/mlverse/tabnet/workflows/R-CMD-check/badge.svg)](https://github.com/mlverse/tabnet/actions) [![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html) [![CRAN status](https://www.r-pkg.org/badges/version/tabnet)](https://CRAN.R-project.org/package=tabnet) [![](https://cranlogs.r-pkg.org/badges/tabnet)](https://cran.r-project.org/package=tabnet) [![Discord](https://img.shields.io/discord/837019024499277855?logo=discord)](https://discord.com/invite/s3D5cKhBkx)

<!-- badges: end -->

An R implementation of: [TabNet: Attentive Interpretable Tabular Learning](https://arxiv.org/abs/1908.07442) [(Sercan O. Arik, Tomas Pfister)](https://doi.org/10.48550/arXiv.1908.07442).\
The code in this repository is an R port using the [torch](https://github.com/mlverse/torch) package of [dreamquark-ai/tabnet](https://github.com/dreamquark-ai/tabnet) PyTorch's implementation.\
TabNet is augmented with [Coherent Hierarchical Multi-label Classification Networks](https://proceedings.neurips.cc//paper/2020/file/6dd4e10e3296fa63738371ec0d5df818-Paper.pdf) [(Eleonora Giunchiglia et Al.)](https://doi.org/10.48550/arXiv.2010.10151) for hierarchical outcomes.

## Installation

Tabnet is temporarily archived on CRAN. We are working hard to get it back.  

In the meantime, you can install the released version from r-universe with:
``` r
install.packages('tabnet', repos = c('https://mlverse.r-universe.dev', 'https://cloud.r-project.org'))
```

The development version can be installed from [GitHub](https://github.com/mlverse/tabnet) with:

``` r
# install.packages("remotes")
remotes::install_github("mlverse/tabnet")
```

## Basic Binary Classification Example

Here we show a **binary classification** example of the `attrition` dataset, using a **recipe** for dataset input specification.

```{r model-fit}
library(tabnet)
suppressPackageStartupMessages(library(recipes))
library(yardstick)
library(ggplot2)
set.seed(1)

data("attrition", package = "modeldata")
test_idx <- sample.int(nrow(attrition), size = 0.2 * nrow(attrition))

train <- attrition[-test_idx,]
test <- attrition[test_idx,]

rec <- recipe(Attrition ~ ., data = train) %>% 
  step_normalize(all_numeric(), -all_outcomes())

fit <- tabnet_fit(rec, train, epochs = 30, valid_split=0.1, learn_rate = 5e-3)
autoplot(fit)
```

The plots gives you an immediate insight about model over-fitting, and if any, the available model checkpoints available before the over-fitting

Keep in mind that **regression** as well as **multi-class classification** are also available, and that you can specify dataset through **data.frame** and **formula** as well. You will find them in the package vignettes.

## Model performance results

As the standard method `predict()` is used, you can rely on your usual metric functions for model performance results. Here we use {yardstick} :

```{r}
metrics <- metric_set(accuracy, precision, recall)
cbind(test, predict(fit, test)) %>% 
  metrics(Attrition, estimate = .pred_class)
  
cbind(test, predict(fit, test, type = "prob")) %>% 
  roc_auc(Attrition, .pred_No)
```

## Explain model on test-set with attention map

TabNet has intrinsic explainability feature through the visualization of attention map, either **aggregated**:

```{r model-explain}
explain <- tabnet_explain(fit, test)
autoplot(explain)
```

or at **each layer** through the `type = "steps"` option:

```{r step-explain}
autoplot(explain, type = "steps")
```

## Self-supervised pretraining

For cases when a consistent part of your dataset has no outcome, TabNet offers a self-supervised training step allowing to model to capture predictors intrinsic features and predictors interactions, upfront the supervised task.

```{r step-pretrain}
pretrain <- tabnet_pretrain(rec, train, epochs = 50, valid_split=0.1, learn_rate = 1e-2)
autoplot(pretrain)
```

The example here is a toy example as the `train` dataset does actually contain outcomes. The vignette on [Self-supervised training and fine-tuning](https://mlverse.github.io/tabnet/articles/selfsupervised_training.html) will gives you the complete correct workflow step-by-step.

## Missing data in predictors

{tabnet} leverage the masking mechanism to deal with missing data, so you don't have to remove the entries in your dataset with some missing values in the predictors variables.

See [`vignette("Missing_data_predictors")`](articles/Missing_data_predictors.html)

# Comparison with other implementations

| Group | Feature | {tabnet} | dreamquark-ai | fast-tabnet |
|--------------|-----------------|:------------:|:------------:|:------------:|
| Input format | data-frame | ✅ | ✅ | ✅ |
|  | formula | ✅ |  |  |
|  | recipe | ✅ |  |  |
|  | Node | ✅ |  |  |
|  | missings in predictor | ✅ |  |  |
| Output format | data-frame | ✅ | ✅ | ✅ |
|  | workflow | ✅ |  |  |
| ML Tasks | self-supervised learning | ✅ | ✅ |  |
|  | classification (binary, multi-class) | ✅ | ✅ | ✅ |
|  | unbalanced binary classification | ✅ |  |  |
|  | regression | ✅ | ✅ | ✅ |
|  | multi-outcome | ✅ | ✅ |  |
|  | hierarchical multi-label classif. | ✅ |  |  |
| Model management | from / to file | ✅ | ✅ | v |
|  | resume from snapshot | ✅ |  |  |
|  | training diagnostic | ✅ |  |  |
| Interpretability |  | ✅ | ✅ | ✅ |
| Performance |  | 1 x | 2 - 4 x |  |
| Code quality | test coverage | 85% |  |  |
|  | continuous integration | 4 OS including GPU |  |  |

: Alternative TabNet implementation features
