---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# tabnet

<!-- badges: start -->
[![R build status](https://github.com/mlverse/tabnet/workflows/R-CMD-check/badge.svg)](https://github.com/mlverse/tabnet/actions)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html)
[![CRAN status](https://www.r-pkg.org/badges/version/tabnet)](https://CRAN.R-project.org/package=tabnet)
[![](https://cranlogs.r-pkg.org/badges/tabnet)](https://cran.r-project.org/package=tabnet)
[![Discord](https://img.shields.io/discord/837019024499277855?logo=discord)](https://discord.com/invite/s3D5cKhBkx)
<!-- badges: end -->

An R implementation of: [TabNet: Attentive Interpretable Tabular Learning](https://arxiv.org/abs/1908.07442).
The code in this repository is an R port of [dreamquark-ai/tabnet](https://github.com/dreamquark-ai/tabnet) PyTorch's implementation using the [torch](https://github.com/mlverse/torch) package.

## Installation

You can install the released version from CRAN with:

```r
install.packages("tabnet")
```

The development version can be installed from [GitHub](https://github.com/mlverse/tabnet) with:

``` r
# install.packages("remotes")
remotes::install_github("mlverse/tabnet")
```

## Basic Example

```{r model-fit}
library(tabnet)
suppressPackageStartupMessages(library(recipes))
library(yardstick)
library(ggplot2)
set.seed(1)

data("attrition", package = "modeldata")
test_idx <- sample.int(nrow(attrition), size = 0.2 * nrow(attrition))

train <- attrition[-test_idx,]
test <- attrition[test_idx,]

rec <- recipe(Attrition ~ ., data = train) %>% 
  step_normalize(all_numeric(), -all_outcomes())

fit <- tabnet_fit(rec, train, epochs = 30, valid_split=0.1, learn_rate = 5e-3)
autoplot(fit)
```

## Model performance results

Model performance results are easy to compute: 

```{r}
metrics <- metric_set(accuracy, precision, recall)
cbind(test, predict(fit, test)) %>% 
  metrics(Attrition, estimate = .pred_class)
  
cbind(test, predict(fit, test, type = "prob")) %>% 
  roc_auc(Attrition, .pred_No)
```

## Explain model on test-set with attention map

TabNet has intrinsic explainability feature through the visualization of attention map, either **aggregated**: 
```{r model-explain}
explain <- tabnet_explain(fit, test)
autoplot(explain)
```

or at **each layer** through the `type = "steps"` option:
```{r step-explain}
autoplot(explain, type = "steps")
```

