---
title: "Interpretation tools"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Interpretation tools}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: 72
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = torch::torch_is_installed(),
  out.width = "100%", 
  out.height = "300px", 
  fig.width = 14
)
```

```{r setup}
library(tabnet)
library(tidyverse, warn.conflicts = FALSE)
set.seed(1)
torch::torch_manual_seed(1)
```

TabNet claims to be an interpretable model because the underlying neural
network architecture uses feature selection masks that can help identify
which features were used at each step.

The paper also defines an aggregate measure that combines the masks of
each step into a single measure.

Note that the selection of important features in the masks is done
instance-wise. Thus, you can identify, for each observation, which
features were considered relevant .

## Experiments

To show how to use the interpretation tools of `tabnet`, we are going to
perform 2 experiments using synthetic datasets very similar to those
that were used in the paper.

### Datasets

First, let's define the functions that we will use to generate data:

-   `make_syn2` will generate a dataset with 10 columns, but only
    columns 3-6 are used to calculate the `y` response vector. This is
    similar to *Syn2* in the paper.

-   `make_syn4` will generate a dataset with 10 columns too. The
    response vector depends on column 10: if the value is greater than
    0, we use columns 1-2 to compute the logits, otherwise we use
    columns 5-6.

```{r}
logit_to_y <- function(logits) {
  p <- exp(logits)/(1 + exp(logits))
  y <- factor(ifelse(p > 0.5, "yes", "no"), levels = c("yes", "no"))
  y
}

make_random_x <- function(n) {
  x <- as.data.frame(lapply(1:10, function(x) rnorm(n)))
  names(x) <- sprintf("V%02d", 1:10)
  x
}

make_syn2 <- function(n = 10000) {
  x <- make_random_x(n)
  logits <- rowSums(x[,3:6])
  x$y <- logit_to_y(logits)
  x
}

make_syn4 <- function(n = 10000) {
  x <- make_random_x(n)
  logits <- ifelse(
    x[,10] > 0,
    rowSums(x[,1:2]),
    rowSums(x[,5:6])
  )
  
  x$y <- logit_to_y(logits)
  x
}
```

Now let's generate the datasets:

```{r}
syn2 <- make_syn2()
syn4 <- make_syn4()
```

### Syn2

Let's fit a TabNet model to the `syn2` dataset and analyze the
interpretation metrics.

```{r}
fit_syn2 <- tabnet_fit(y ~ ., syn2, epochs = 10, verbose = TRUE)
```

In the feature importance plot we can see that, as expected, features
`V03-V06` are by far the most important ones.

```{r}
vip::vip(fit_syn2)
```

Now let's visualize the aggregated-masks plot. In this figure we see
each observation on the x axis and each variable on the y axis. The
colors represent the importance of the feature in predicting the value
for each observation.

```{r}
library(tidyverse)
ex_syn2 <- tabnet_explain(fit_syn2, syn2)

autoplot(ex_syn2)
```

We can see that the region between V03 and V06 concentrates most of the
higher intensity colors, and the other variables are close to 0. This is
expected because those are the variables that we considered when
building the dataset.

Next, we can visualize the attention masks for each step in the
architecture.

```{r}
autoplot(ex_syn2, type="steps")
```

We see that the first step captures a lot of noise, but the other 2
steps focus specifically on the important features.

## Syn 4

Now let's analyze the results for the Syn4 dataset. This dataset is a
little more complicated for TabNet because there's a strong interaction
between the variables. Depending on V10, different variables are used to
create the response variable and we expect to see this in the masks.

First we fit the model for 10 epochs.

```{r}
fit_syn4 <- tabnet_fit(y ~ ., syn4, epochs = 10, verbose = TRUE)
```

In the feature importance plot we have, as expected, strong importance
for `V10`, and the other features that are used conditionally - either
`V01-V02` or `V05-V06`.

```{r}
vip::vip(fit_syn4)
```

Now let's visualize the attention masks. Notice that we arranged the
dataset by `V10` so we can easily visualize the interaction effects.

We also trimmed to the 99th percentile so the colors shows the
importance even if there are strong outliers.

```{r}
ex_syn4 <- tabnet_explain(fit_syn4, arrange(syn4, V10))

autoplot(ex_syn4, quantile=.995)
```

From the figure we see that V10 is important for all observations. We
also see that for the first half of the dataset `V05` and `V06` is the
most important feature, while for the other half, `V01` and `V02` are
the important ones.

We can also visualize the masks at each step in the architecture.

```{r}
autoplot(ex_syn4, type="steps", quantile=.995)
```

We see that step 1 and 3 both focus on `V10`, but on different
additional features, depending on `V10`. Step 2 seems to have found some
noise in `V08`, but also focuses strongly on `V01-V02` and `V05-V06`.
