---
title: "Fitting tabnet with tidymodels"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Fitting tabnet with tidymodels}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

```{r setup}
library(tabnet)
library(tidymodels)
library(modeldata)
```

In this vignette we show how to create a TabNet model using the tidymodels interface.

We are going to use the `lending_club` dataset available
in the `modeldata` package.

First let's split our dataset into training and testing so we can later access performance of our model:

```{r}
set.seed(123)
data("lending_club", package = "modeldata")
split <- initial_split(lending_club, strata = Class)
train <- training(split)
test  <- testing(split)
```

We now define our pre-processing steps. Note that tabnet handles categorical variables, so we don't need to do any kind of transformation to them. Normalizing the numeric variables is a good idea though.

```{r}
rec <- recipe(Class ~ ., train) %>%
  step_normalize(all_numeric())
```

Next, we define our model. We are going to train for 50 epochs with a batch size of 128. There are other hyperparameters but, we are going to use the defaults.

```{r}
mod <- tabnet(epochs = 50, batch_size = 128) %>%
  set_engine("torch", verbose = TRUE) %>%
  set_mode("classification")
```

We also define our `workflow` object:

```{r}
wf <- workflow() %>%
  add_model(mod) %>%
  add_recipe(rec)
```

We can now define our cross-validation strategy:

```{r}
folds <- vfold_cv(train, v = 5)
```

And finally, fit the model:

```{r}
fit_rs <- wf %>%
  fit_resamples(folds)
```

After a few minutes we can get the results:

```{r}
collect_metrics(fit_rs)
```

```
# A tibble: 2 x 5
  .metric  .estimator  mean     n  std_err
  <chr>    <chr>      <dbl> <int>    <dbl>
1 accuracy binary     0.946     5 0.000713
2 roc_auc  binary     0.732     5 0.00539 
```

And finally, we can verify the results in our test set:

```{r}
model <- wf %>% fit(train)
test %>% 
  bind_cols(
    predict(model, test, type = "prob")
  ) %>% 
  roc_auc(Class, .pred_bad)
```

```
# A tibble: 1 x 3
  .metric .estimator .estimate
  <chr>   <chr>          <dbl>
1 roc_auc binary         0.710
```





